I do not spend much time watching reality television, but the Discovery Channel's Gold Rush is an exception.
When my wife asks me to turn it off, I explain to her that Gold Rush is helping me finish my PhD.
I am not sure she ever bought it, but I can definitely draw some similarities between the show and my pursuit of a doctoral degree.
Mining happens on two kinds of claims.
Some claims have already been mined.
The nuggets have been taken.
However, the good areas are known and new tools allow for the extraction of gold flakes left behind.

The second kind of claim is known as \emph{virgin ground}.
This area has never been mined, so gold nuggets can potentially be found.
My advisor, Jeff Shamma, showed me some \emph{virgin ground} at the onset of my PhD and for this I will forever be thankful.

On \emph{virgin ground}, you spend days, months, and in my case, years, taking the lay of the land.
In both the pursuit of gold and the PhD, the unexplored territories are daunting.
However, regardless how painful the process, finding gold nuggets feels like magic.
Jeff, you gave me the claim and stayed by my side until I struck gold.
You never hovered over my shoulder, but you continually guided me towards the treasure.
Your ability to help take a step back and think about the core of a problem is amazing.
Thank you for seeing me through this opportunity of a lifetime.
I know sometimes I was a discouraged miner, but you helped me keep my chin up.
It was truly a pleasure working with you.

Secondly, I would like to thank my co-advisor, Éric Féron.
Your continued support and industry-oriented approach was the perfect complement to my theoretical thesis.
Even though I did not sit in your lab, you always were inclusive and helped me feel like I was part of your team.

Advisors guide you through a PhD, but labmates are there to catch you when you fall.
Without them, it would oftentimes be hard to get back up.
For those who have been through the process, you understand that getting a PhD takes team work.
Thank you for being there and helping me make my way through the mine.

Next, I want to thank those who became close confidants during my doctoral studies.
Some of you were in my lab, others of you I met through classes and symposiums.
Most of the time we spoke at Starbucks over doppio espressos.
Thank you to Francesco Barale, Romain Pelard, Patrick Melet, Travis Deyle, Pierrick Burgain, William Mantzel, Michael Fox, Malachi Jones, Georgios Kotsalis, Georgios Chasparis, Georgios Piliouras, Aurèle Balavoine, Chris Turnes, Karen Pottin, Sam Shapero, Taylor Shapero, Vlad Popescu, Romain Jobredeaux, Aude Marzuoli, Emmanuel Boidot, Florian Hauer, Tim Wang, Ibrahim al-Shyoukh, Yasin Yazicioglu, Daniel Pickem, Rowland O'Flaherty, Greg Droge, Jean-Pierre de la Croix, Baptiste Coudriller, Behrouz Touri, Sebastian Ruff, Lichun Li, and Kwang-Ki Kim.

I also want to thank my committee members, Fumin Zhang, Spyros Reveliotis, Nina Balcan, and Yorai Wardi, for their invaluable feedback.
Sometimes you can get tunnel vision when you are digging for gold.
However, your insight helped me take a step back and focus on the big picture.
I appreciate your help in letting me chisel the dirt away from my dissertation so it would have a chance to sparkle.

My acknowledgment section would not be complete without recognizing my family.
My parents, grandparents, mother-in-law, and siblings were always supportive of me pursuing a doctoral degree.
I cannot imagine this journey without their love and guidance.
Lastly, I must thank my wife for her many sacrifices so that I could strike gold.
Laura was always there for me.
She listened when I needed to talk and texted me during the day when I needed encouragement.
Laura, thank you for your support.
You deserve every bit of your PhT because you really did, \emph{put hubby through}.
